Reproducible Research: Peer Assessment 1
==========================
## Loading and Preprocessing the data  
```{r,results="hide"}
data = read.csv("activity.csv")
data1 = data[complete.cases(data),]
```

## Histogram of the total number of steps in each day  
```{r}
library(ggplot2)
library(dplyr)
data$date = as.Date(data$date)
data1$date = as.Date(data1$date)
y = tapply(data1$steps,data1$date,sum)
x = unique(data1$date)
ques_1 = data.frame("date" = x,"Total" = y)
z = rep(ques_1$date,ques_1$Total)
z = as.data.frame(z)
names(z) = "Date"
f = ggplot(z,aes(x = Date))
f+geom_histogram(binwidth=1)+ggtitle("Total Steps")
```

## Mean and Median number of steps  
The following table shows the mean and median of the total number of steps in each day.  
```{r,results = "asis"}
summaryy = data1%>%group_by(date)%>%summarise(mean = mean(steps),median = median(steps))
summaryy = data.frame(summaryy)
library(xtable)
summaryy$date = as.character(summaryy$date)
tab = xtable(summaryy)
print(tab,type = "html")
```

## Average number of steps taken  
```{r}
data2 = data1
avg = data2%>%group_by(interval)%>%summarise(average = mean(steps))
with(avg,plot(interval,average,type = "l",ylab = "Average Steps"))
```
  
## The interval with maximum number of steps  
```{r}
index = which.max(avg$average)
max_interval = avg[index,1]
i = max_interval[[1]]
i
```
The interval with maximum number of steps is **`r i`**.

## Imputing the missing data  
The missing values in each interval is replaced with the average number of steps taken in that interval averaged over the total days.  
```{r}
total_missing = sum(is.na(data$steps))
total_missing
filled_data = data
for(i in 1:nrow(data)){
        if(is.na(filled_data[i,1])){
                c = avg$interval%in%filled_data[i,3]
                filled_data[i,1] = avg[c,2]
        }
}
```

## Histogram of the total number of steps taken each day after missing values are imputed  
```{r}
y_new = tapply(filled_data$steps,filled_data$date,sum)
x_new = unique(filled_data$date)
ques_3 = data.frame("date" = x_new,"Total" = y_new)
z_new = rep(ques_3$date,round(ques_3$Total))
z_new = as.data.frame(z_new)
names(z_new) = "Date"
f2 = ggplot(z_new,aes(x = Date))
f2+geom_histogram(binwidth=1)+ggtitle("Total Steps")
```

## Mean and Median after imputing the values  
The following table shows the mean and the median of the total number of steps of each day afte imputing the missing values.  
```{r,results = "asis"}
summaryy2 = filled_data%>%group_by(date)%>%summarise(mean = mean(steps),median = median(steps))
summaryy2 = data.frame(summaryy2)
summaryy2$date = as.character(summaryy2$date)
tab = xtable(summaryy2)
print(tab,type = "html")
```
  
**Since the missing values corresponded to one whole single day. Imputing the values doesn't have any effect on the mean and the median of the other days.**

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends  
```{r}
filled_data$day = factor(weekdays(filled_data$date)%in%c("Saturday","Sunday"),labels = c("Weekdays","weekends"))
avg2 = filled_data%>%group_by(interval,day)%>%summarise(average = mean(steps))
g = ggplot(avg2,aes(interval,average))
g + geom_line()+facet_grid(day~.)
```
  
**This panel plot clearly shows that the number of steps taken on weekdays is higher than the number of steps taken on weekends.**