---
output:
  pdf_document: default
  html_document: default
---
Effect or transmission on MPG
=======================================

## Executive Summary  

```{r,echo = F,results='hide',include=F}
data = mtcars
library(GGally)
data$am = factor(mtcars$am,labels = c("automatic","manual"))
f3 = lm(mpg~disp+am+wt+hp,data = data)
```

- Overall, this analysis suggests that the mpg for **manual trannsmission is better than that of the automatic transmission**.  
- It suggests that keeping the other covariates constant, the mean mpg for manual transmission is greater than that of automatc transmission by *`r summary(f3)$coefficients[3,1]`* with the p-value of  *`r summary(f3)$coefficients[3,4]`*.  
- This behavior can also be shown using the two sided t test as shown below.    

```{r,echo = F}
t.test(mpg~am,data = data,paired = F)$conf
```
 
The confidence interval lies entirely below zero suggesting that mpg for manual transmission is better than that of the automatic transmission.  

## Basic Exploratory Analysis   

```{r,echo=F,fig.height=3,fig.width=6}
library(ggplot2)
data = mtcars
data$am = factor(mtcars$am,labels = c("automatic","manual"))
g = ggplot(data = data,aes(y = mpg,x = am,fill = am))
g+geom_boxplot()
```

This plot gives a rough estimate that the mpg for manual is better than that of automatic transmission.  
Let's dig into it and try to fit a linear regression model which may give a better picture of it.  
We also find that:  
- cyl,gear and carb are discrete variables.  
- vs and am are factor variables.  
-  mpg,disp,hp,drat,wt and qsec are continuous variables.  

## Model Selection

```{r,echo = F}
f = lm(mpg~.,data = data)
hatvalues(f)
```

The hatvalues suggests that there is no such potential influencer in our data set that we may need to look upon.  

Let's fit the model with different parameters depending upon their correlation with the mpg and see which one fits the best.   

```{r,echo = F}
f1 = lm(mpg~disp,data = data)
f1 = lm(mpg~disp+am,data = data)
f2 = lm(mpg~disp+am+wt,data = data)
f3 = lm(mpg~disp+am+wt+hp,data = data)
f4 = lm(mpg~disp+am+wt+hp+drat,data = data)
f5 = lm(mpg~disp+am+wt+hp+drat+qsec,data = data)
anova(f1,f2,f3,f4,f5)
```

As we can see that model 2 performs better than the first model and the model 3 perfoms even better than the model 2 with the F value of apprx. 0.0026.  
After that the F value increases which suggests that the later models are not good.    
This analayis of variance suggests that the third model is best among all of them.   
  
## Analysing  

```{r,echo = F,fig.height=3,fig.width=6}
d = cbind(predict(f3),resid(f3),data$am)
d = data.frame(d)
g = ggplot(d,aes(X1,X2))
g+geom_point(aes(color = X3))+ggtitle("Residuals vs Fitted")+labs(x = "fitted",y = "residuals")
```

There is no such significant pattern in the residuals and the fitted values as such.  

```{r,echo = F}
summary(f3)$coefficients
```

The coefficient corresponding to "ammanual" suggests that keeping the other covariates constant, the **mean mpg for manual transmission is greater than that of automatc transmission** by `r summary(f3)$coefficients[3,1]` with the p-value of  `r summary(f3)$coefficients[3,4]`.  
Also the coefficient corresponding to displ suggests that there is a very *minor increase in the mpg with displacement* keeping other covariates constant.  
The coefficient corresponding to weight suggests that *wt has a negative impact on the the mpg* keeping others constant.  
Also, there is a very *minor decrease in the mpg with hp* keeping other covariates constant.  

## Appendix Figures  
```{r,cache = T,echo = F}
data$vs = factor(data$vs)
g = ggpairs(data)
g
par(mfrow = c(2,2))
plot(f3)
```
  